\tool: Automated Generation of Diversified Tests for Self-driving Cars from Existing Maps

### 2021-08-20 paper/main.tex
Finished pass on section 4

\newcommand{\tool}{\textsc{SALVO}\xspace}
\newcommand{\challenge}{2021 IEEE Autonomous Driving AI Test Challenge\xspace}
\newboolean{showcomments}
\setboolean{showcomments}{true}         
%\setboolean{showcomments}{false} 
\ifthenelse{\boolean{showcomments}}
  {\newcommand{\nb}[2]{
  \fbox{\bfseries\sffamily\scriptsize#1}
     {\sf\small$\blacktriangleright$\textit{\textcolor{red}{#2}}$\blacktriangleleft$}
   }
  }
  {\newcommand{\nb}[2]{}
   \newcommand{\cvsversion}{}
  }

@@

\begin{abstract}
% Testing self-driving car software using photo-realistic and accurate physical simulators is becoming a widely adopted practice. Simulators enable the automated execution of many tests that possibly implement safety-critical driving scenarios that are too dangerous to put into practice. However, despite the speed-up achieved by using simulators, fundamental challenges are yet to be addressed before achieving cost-effective testing of self-driving cars.
% %
% In this paper, we present \tool, an approach to automated self-driving car software testing that addresses some of those challenges. Specifically, \tool
% \begin{inparaenum}[(i)]
% \item systematically identifies relevant driving scenarios among a myriad of potential driving scenarios that can be defined over existing maps;
% \item selects driving scenarios that reduce testing efforts while ensuring quantifiable different behaviors, and 
% \item automatically generates test cases that implement those diversified driving scenarios.
% \end{inparaenum}
% %
% We demonstrated \tool in the \challenge
% by testing state-of-the-art self-driving car software using an industrial driving simulator.
% In a matter of minutes, \tool analyzed five maps of different complexity levels (from $2$ to more than $15$ intersections), identified more than $1100$ relevant driving scenarios, and selected on average half of them for the execution.
% %
% The evaluation results showed that tests cases generated by \tool from the selected driving scenarios resulted in quantifiable different trajectories followed by the test subject and exposed issues in it.

%For testing self-driving cars simulations are heavily utilized since these allow running many simulations at once and allow enforcing critical situations without endangering any other drivers or pedestrians.
%For increasing the quality of tests photo-realistic maps and accurate physical simulators are a widely adopted and still growing practice.
%However, despite these advantages there are still fundamental challenges including cost-effectiveness and further increasing the degree of automation for finding or generating critical tests.\\
%This paper presents \tool, an approach
%(1) to systematically identify relevant driving scenarios among the myriad of potential driving scenarios,
%(2) to select scenarios that reduce the required overall testing efforts by ensuring quantifiability and diversity of the expected behaviours of self-driving cars for succeeding scenarios and
%(3) to generate those test cases.\\
%We demonstrated \tool at the \challenge by running state-of-the-art self-driving car software in an industrial driving simulator against our generated test cases.
%In a matter of minutes, \tool analyzed five maps of different complexity (from 2 to more than 15 intersections), identified more than 1100 relevant driving scenarios and selected on average half of them for actual test executions.
%The evaluation results showed that the generated test cases resulted in quantifiable and diverse trajectories and exposed issues of the self-driving car software.

Simulation-based tests are more cost-effective and less dangerous than field tests; hence, they are becoming the norm for thoroughly testing self-driving cars in virtual environments. High-quality simulation-based testing requires physically accurate computer simulations,  detailed and photo-realistic maps, and systematic approaches for generating tests. Moreover, since creating detailed maps is a manual process, they are expensive, and testers should avoid wasting such valuable resources, for example, by generating irrelevant test cases. To address those issues, we propose \tool a fully automated approach to identify quantifiably diverse and critical driving scenarios that can be instantiated on existing high-definition maps and implement them in an industrial driving simulator as executable test cases.
The evaluation of \tool in the context of the \challenge showed that it could analyze maps of different complexity and identify many critical driving scenarios in minutes. Furthermore, the tests \tool generated stressed a state-of-art self-driving car software in quantifiably diverse ways and exposed issues with its implementation.
\end{abstract}

@@

\subsection{Selecting Diversified Scenarios}
Up to this point, \tool does not take the diversity of scenarios into account as it simply generates as many scenarios as possible to explore the possibilities offered by the input road map. However, many legal trajectories may stress the ego-vehicle similarly; for example, they may force it to drive straight through the same intersections from multiple directions. Therefore, \tool filters out test cases that are too similar to each other to improve testing cost-efficiency.

We experimented with two (greedy) approaches for filtering out similar test cases based on distance metrics and feature maps.

\subsubsection{Filtering Similar Scenarios by Similarity}
The first approach to select diversified scenarios is based on computing the similarity of the ego-vehicle' trajectories imposed by the driving scenarios. The chosen metric is Levenshtein distance because it can provide a numerical representation of how different two trajectories are by measuring how many \emph{edits} are required to change one trajectory into the other. Since each trajectory is an array of points, so an \emph{edit} could be an insertion, a deletion, or a substitution of a point. In theory, Levenshtein distance is calculated by initializing an  $m\times n$ matrix where each cell contains the Levenshtein distance between the first $m^{th}$ points of one trajectory and the first $n^{th}$ points of other. The goal is to compute and fill out the entire matrix from the upper left to the lower right corner. Following this way, the number in the bottom-right corner will be Levenshtein distance between both trajectories.
%

Specifically, \tool calculates the iterative Levenshtein distance between pairs of trajectories and discards driving scenarios whose trajectories are not distant enough from (i.e., they are too similar to) the trajectories of the driving scenarios already selected (see Figure~\ref{fig:similarity}).
%
Testers can control the similarity between the trajectories using a configurable threshold to create larger or smaller test suites.

A similar approach to select diverse trajectories was followed by Zohdinasab and co-author in a pilot study to identify road features relevant for testing lane keeping systems~\cite{DBLP:conf/issta/ZohdinasabRGT21}.
As \tool, they took inspiration for the computation of the iterative Levenshtein distance from the work of Riccio and Tonella~\cite{DBLP:conf/sigsoft/RiccioT20}.

\subsubsection{Filtering Driving Scenarios using Feature Maps}
The second approach to select diversified scenarios is based on the feature maps defined by Zohdinasab et al.~\cite{DBLP:conf/issta/ZohdinasabRGT21}. 
%
\tool generates numerous driving scenarios containing different types of trajectories and features such as minimum curvature and direction coverage capturing high-level characteristics of those trajectories, which can be used to build feature maps. Those maps place driving scenarios in a two-dimensional space according to the values taken by their features.
%
Intuitively, trajectories with similar features are placed in nearby cells, whereas trajectories with different features are placed in distance cells (see Figure~\ref{fig:feature-maps}).

\begin{figure*}[t]
\minipage{0.32\textwidth}
\includegraphics[width=\linewidth]{images/feature_cubetown}
\endminipage\hfill
\minipage{0.32\textwidth}
\includegraphics[width=\linewidth]{images/two_roads_same_cell.png}
\endminipage\hfill
\minipage{0.32\textwidth}%
  \includegraphics[width=\linewidth]{images/roads_different_cell.png}
\endminipage
\caption{Left: Example of feature map distance coverage (dc) $\times$ min radius (mr) generated by \tool from the Cube Town map. Central: Trajectories sampled from the same cell in the feature map. Right: Trajectories sampled from distance cells in the feature map.}
\label{fig:feature-maps}
\end{figure*}

Given a feature map, \tool considers similar those trajectories that are placed in the same cells. Consequently, it selects from each cell at most one trajectory (i.e., driving scenario) as representative, while discards all the others.

Notably using the feature maps, testers can easily quantify the coverage achieved by \tool by counting the number of cells that contain at least one driving scenario. Likewise, developers can use the feature maps to assess the "richness" of the input road maps (see Figure~\ref{fig:feature-maps-results}).

% \begin{figure}[H]
%   \centering
%     \includegraphics[width=0.5\textwidth]{images/feature_borregasave_maps}
%     \caption{Borregas Ave trajectories based 10-by-10 cells map.}
% \end{figure}

% \begin{figure}[H]
%   \centering
%     \includegraphics[width=0.5\textwidth]{images/two_roads_same_cell.png}
%     \caption{Example of two roads that belong to the same cell.}
% \end{figure}

% \begin{figure}[H]
%   \centering
%     \includegraphics[width=0.5\textwidth]{images/roads_different_cell.png}
%     \caption{Example of three roads taken from different (distant) cells.}
% \end{figure}

\subsection{Generating Test Cases}
\tool generates a set of diversified driving scenarios as a combination of lanelets. So it needs to transform them into test cases that can be executed as simulations and result either in passing or failing.
\tool does so by implementing introductory test oracles that check if the ego-vehicle reaches the target location after the intersection within a predefined timeout and by using simple geometric transformations to place the ego-vehicle in the starting position and rotate it according to the geometry of the initial lanelets.

However, doing so is not enough since the simulations may include additional elements, such as traffic lights that affect the test result. For example, if the ego-vehicle needs to pass a controlled intersection and the traffic light is red, the ego-vehicle must stop in front of it, which may cause the timeout to trigger a false positive. Therefore, to avoid such false positives, \tool programs any traffic lights on the expected trajectories to turn green when the ego-vehicle approaches them.

\subsection{Spicing Up the Test Cases}
\label{sec:spicing-up}
The tests generated by \tool are plain: they stress the ability of the ego-vehicle to plan trajectories that pass intersections and keep the lane but involve no other element besides the ego-vehicle and possibly traffic lights. 
Therefore, we illustrate how those test cases can be extended by placing static obstacles on the course of the ego-vehicle by letting  \tool place parked vehicles on the roads.

Our approach to ``spice up'' the test cases starts from an existing driving scenario and generate tests by placing NPC vehicles predefined positions between the ego-vehicle's initial position (i.e., the starting point of the driving scenario) and the intersection. Placing NPC vehicles results is partially or entirely occlusions of the lane, forcing the ego-vehicle to swerve or change lane. Default occlusions are partial, and \tool causes them by placing an NPC vehicle on the left or right side of the lane. Additionally, if the geometry of the intersection allows it, \tool places an NPC in the middle of the lane in front of the ego-vehicle, forcing it to switch lanes before entering the intersection.

In summary, on top of the test cases that \tool can generate from a given map, testers can quickly generate up to four times more test cases by simply adding NPC vehicles in front of the ego-vehicle. As we did before, we decided to place up to one NPC vehicle per test case to keep the test cases as simple and understandable as possible.

\subsection{Prototype}
\tool is implemented as a Python (version 3.9) command-line application and currently integrated with the SVL driving simulator.\footnote{\href{https://www.svlsimulator.com/}{https://www.svlsimulator.com/}} 
% Specifically, we use Python 3.9, the libraries listed in the requirements.txt file, and the maps in OpenDrive format available in the SVL simulator.
%
We tested \tool on Ubuntu 20.04.2 LTS running the SVL Simulator version 2021.01 and Apollo Baidu version 6.0 \footnote{\href{https://apollo.auto/}{https://apollo.auto/}} in accordance to the rules of the \challenge.

\section{Achieved Results and Found Problems}
To evaluate \tool, we applied it to test Apollo Baidu on the following five maps that were available on the SVL Web site at the time the \challenge took place:

@@

\begin{figure*}[!tp]
\minipage{0.31\textwidth}
  \includegraphics[width=\linewidth]{images/map_cubetown}
  \includegraphics[width=\linewidth]{images/feature_cubetown}
  \caption{Cube Town map.}
\endminipage\hfill
\minipage{0.31\textwidth}
  \includegraphics[width=\linewidth]{images/map_borregasave}
  \includegraphics[width=\linewidth]{images/feature_borregasave}
  \caption{Borregas Ave map.}
\endminipage\hfill
\minipage{0.31\textwidth}%
  \includegraphics[width=\linewidth]{images/map_shalun}
  \includegraphics[width=\linewidth]{images/feature_shalun}
  \caption{Shalun map.}
\endminipage

\minipage{0.31\textwidth}%
  \includegraphics[width=\linewidth]{images/map_gomentum}
  \includegraphics[width=\linewidth]{images/feature_gomentum}
  \caption{Gomentum map.}
\endminipage\hfill
\minipage{0.31\textwidth}%
  \includegraphics[width=\linewidth]{images/map_sanfrancisco.png}
  \includegraphics[width=\linewidth]{images/feature_sanfrancisco.png}
  \caption{San Francisco map.}
\endminipage
\minipage{0.31\textwidth}%
    \includegraphics[width=\linewidth]{images/map_sanfrancisco.png}
\endminipage
\label{fig:feature-maps-results}
\end{figure*}

>>>>

\newcommand{\tool}{\textsc{SALVO}\xspace}
\newcommand{\challenge}{2021 IEEE Autonomous Driving AI Test Challenge\xspace}

\newboolean{showcomments}
\setboolean{showcomments}{true}         
%\setboolean{showcomments}{false} 
\ifthenelse{\boolean{showcomments}}
  {\newcommand{\nb}[2]{
  \fbox{\bfseries\sffamily\scriptsize#1}
     {\sf\small$\blacktriangleright$\textit{\textcolor{red}{#2}}$\blacktriangleleft$}
   }
  }
  {\newcommand{\nb}[2]{}
   \newcommand{\cvsversion}{}
  }

@@

\begin{abstract}
Simulation-based tests are more cost-effective and less dangerous than field tests; hence, they are becoming the norm for thoroughly testing self-driving cars in virtual environments. High-quality simulation-based testing requires physically accurate computer simulations,  detailed and photo-realistic maps, and systematic approaches for generating tests. Moreover, since creating detailed maps is a manual process, they are expensive, and testers should avoid wasting such valuable resources, for example, by generating irrelevant test cases. To address those issues, we propose \tool a fully automated approach to identify quantifiably diverse and critical driving scenarios that can be instantiated on existing high-definition maps and implement them in an industrial driving simulator as executable test cases.
The evaluation of \tool in the context of the \challenge showed that it could analyze maps of different complexity and identify many critical driving scenarios in minutes. Furthermore, the tests \tool generated stressed a state-of-art self-driving car software in quantifiably diverse ways and exposed issues with its implementation.
\end{abstract}

@@

\subsection{Selecting Diversified Scenarios}
Up to this point, \tool does not take the diversity of scenarios into account as it simply generates as many scenarios as possible to explore the possibilities offered by the input road map. However, many legal trajectories may stress the ego-vehicle similarly; for example, they may force it to drive straight through the same intersections from multiple directions. Therefore, \tool filters out test cases that are too similar to each other to improve testing cost-efficiency.
%
We experimented with two (greedy) approaches for filtering out test cases based on distance metrics and feature maps.

\subsubsection{Filtering Similar Scenarios by Similarity}
The first approach to select diversified scenarios relies on the similarity between the trajectories imposed by the driving scenarios on ego-vehicle. After an initial exploration considering standard metrics such as edit distance and cosine similarity (see~Figure~\ref{fig:similarity}), we decided to adopt the Levenshtein distance to quantify the similarity of generated scenarios. The Levenshtein distance quantifies the similarity between trajectories by measuring how many \emph{edits} are required to transform one trajectory into the other; the smaller the number of edits is, the more similar the two trajectories are. Notably, \tool can use the Levenshtein distance because we represent trajectories as coordinates arrays, and edits correspond to insertions, deletions, and substitutions of arrays' entries. Specifically, \tool calculates the Levenshtein distance between pairs of trajectories using the code provided by DeepJanus~\cite{DBLP:conf/sigsoft/RiccioT20} and discards all the driving scenarios whose trajectories are too similar to the trajectories of driving scenarios already selected. 
%
Testers can control the similarity between the trajectories using a configurable threshold to create larger or smaller test suites. For completeness, in our evaluation we filtered out trajectories with a distance value smaller than $1.9$.

%A similar approach to select diverse trajectories was followed by Zohdinasab and co-author in a pilot study to identify road features relevant for testing lane keeping systems~\cite{DBLP:conf/issta/ZohdinasabRGT21}.
%As \tool, they took inspiration for the computation of the iterative Levenshtein distance from the work of Riccio and Tonella~\cite{DBLP:conf/sigsoft/RiccioT20}.

\subsubsection{Filtering Driving Scenarios using Feature Maps}
The second approach to select diversified scenarios utilizes the feature maps defined by Zohdinasab et al.~\cite{zohdinasab2021deephyperion}. 
Trajectory features capture high-level characteristics of the abstract scenarios that are relevant for testing self-driving car software; examples of those features are the number of turns and the minimum curvature\footnote{The minimum curvature highlights the presence of sharp turns.} that quantify how difficult it would be to follow the trajectory. Instead, feature maps place driving scenarios in a two-dimensional space according to the values taken by their features and provide a compact and understandable representation of their most salient aspects. Intuitively, scenarios with similar features are placed in nearby cells, whereas scenarios that are far away in the map have very different features (see Figure~\ref{fig:feature-maps}).

\begin{figure*}[t]
\minipage{0.33\textwidth}
\includegraphics[width=\linewidth]{images/feature_cubetown}
\endminipage\hfill
\minipage{0.33\textwidth}
\includegraphics[width=\linewidth]{images/two_roads_same_cell.png}
\endminipage\hfill
\minipage{0.33\textwidth}%
  \includegraphics[width=\linewidth]{images/roads_different_cell.png}
\endminipage
\caption{Examples of feature map and trajectories for the Cube Town map. 
Left panel: The feature map generated by \tool using direction coverage (dc) and minimum radius (mr) as features.
Central panel: Similar trajectories sampled from the same cell in the feature map, i.e., $(6,1)$.  
Right panel: Diverse trajectories sampled from distant cells in the feature map, i.e., $(10,2), (1,10), (6,1)$.}
\label{fig:feature-maps}
\end{figure*}

\tool's goal is to generate driving scenarios that result in trajectories with different characteristics; therefore, given a feature map filled with all the generated trajectories, \tool selects from each cell at most one trajectory (i.e., driving scenario) as a representative for the class of trajectories with the same feature combination while discarding the others.

Notably, using the feature maps, testers can easily quantify the coverage achieved by \tool by counting the number of cells that contain at least one driving scenario. Likewise, content generators can use the feature maps to assess the ``richness'' of the road maps. 
%
For example, figures~\ref{fig:feature-maps-cubetown}~--~\ref{fig:feature-maps-sanfrancisco} show the road map (top plot) and the corresponding feature map (bottom plot) for the five maps we used in the \challenge (see Section~\ref{sec:evaluation}). From those figures, we can make the following observations: 
\begin{inparaenum}[(1)]
\item road maps of increasing complexity, i.e., maps that contain more intersections, result in feature maps that are more covered, 
and \item road maps with a repetitive structure result in feature maps with very dark cells, i.e., cells that contain many trajectories with very similar features.
\end{inparaenum}

\subsection{Generating Test Cases}
\tool defines abstract driving scenarios as combinations of lanelets. So it needs to concretize them into executable test cases featuring test oracles that check if the ego-vehicle reaches the target location and if it does so within a predefined timeout. Also, \tool must configure the driving simulation to place the ego-vehicle in the starting position and rotate it according to the road direction. However, doing so may not be enough to ensure that the generated tests are reliable and reproducible because the simulations include additional components with randomized behaviors that can affect the test results. For example, if the ego-vehicle needs to pass an intersection controlled by a traffic light, its state affects the behavior of the ego-vehicle: if the traffic light is red, the ego-vehicle must stop before it, which may cause the timeout to trigger a false positive. However, executing multiple times the same test may produce different results, hence flakiness, as the traffic light's initial state is initialized at random unless configured explicitly by the test. Therefore, to avoid false positives and reduce test flakiness, \tool programs the traffic lights on the expected trajectories to turn green when the ego-vehicle approaches them.

\subsection{Spicing Up the Test Cases}
\label{sec:spicing-up}
The tests generated by \tool are diverse but plain: they stress the ability of the ego-vehicle to plan various trajectories to pass the intersections while keeping the lane but involve no other elements besides the ego-vehicle and the traffic lights. To further increase the diversity of the generated tests and enable testing additional features of the ego-vehicle such as object detection and collision avoidance, we let \tool extend the basic tests by placing static obstacles, i.e., NPC vehicles, on the course of the ego-vehicle.

Our approach to ``spice up'' tests consists in placing an NPC vehicle in predefined positions between the ego-vehicle's initial position (i.e., the starting point of the driving scenario) and the intersection to pass such that the NPC vehicle partially or entirely occludes the lane that the ego-vehicles is supposed to keep. Default occlusions are partial and cause the ego-vehicle to swerve before entering the intersection; \tool can place the NPC vehicle either on the left or right side of the lane, but not both. Additionally, but only if the geometry of the intersection allows it, \tool can place the NPC in the middle of the lane in front of the ego-vehicle, forcing it to switch lanes before entering the intersection.

In summary, on top of the diversified test cases that \tool generates from a given map, testers can quickly generate up to four times more test cases by simply letting \tool add NPC vehicles in front of the ego-vehicle. Notably, we limit \tool to place a single NPC vehicle per test case to keep them as simple and understandable as possible.

\subsection{Prototype}
We implemented \tool in Python (v3.9) as a command-line application and integrated it with the SVL driving simulator (v2021.01) in accordance to the rules of the \challenge.
For the evaluation, we ran \tool on a gaming laptop running Ubuntu~20.04.2~LTS, the SVL driving Simulator, and Apollo Baidu (v6.0).

\section{Achieved Results and Problems Found}
\label{sec:evaluation}

To evaluate \tool, we applied it to test Apollo Baidu on the following five maps that were available on the SVL Web site at the time the \challenge took place:

@@

\begin{figure*}[!tp]
\minipage{0.31\textwidth}
  \includegraphics[width=\linewidth]{images/map_cubetown}
  \includegraphics[width=\linewidth]{images/feature_cubetown}
  \caption{Cube Town map.}
  \label{fig:feature-maps-cubetown}
\endminipage\hfill
\minipage{0.31\textwidth}
  \includegraphics[width=\linewidth]{images/map_borregasave}
  \includegraphics[width=\linewidth]{images/feature_borregasave}
  \caption{Borregas Ave map.}
    \label{fig:feature-maps-borregas}
\endminipage\hfill
\minipage{0.31\textwidth}%
  \includegraphics[width=\linewidth]{images/map_shalun}
  \includegraphics[width=\linewidth]{images/feature_shalun}
  \caption{Shalun map.}
    \label{fig:feature-maps-shalun}
\endminipage 
\end{figure*}

\begin{figure*}[!tp]
\hfill
\minipage{0.31\textwidth}%
  \includegraphics[width=\linewidth]{images/map_gomentum}
  \includegraphics[width=\linewidth]{images/feature_gomentum}
  \caption{Gomentum map.}
    \label{fig:feature-maps-gomentum}
\endminipage\hfill
\minipage{0.31\textwidth}%
  \includegraphics[width=\linewidth]{images/map_sanfrancisco.png}
  \includegraphics[width=\linewidth]{images/feature_sanfrancisco.png}
  \caption{San Francisco map.}
    \label{fig:feature-maps-sanfrancisco}
\endminipage
\hfill
%\minipage{0.31\textwidth}%
%    \includegraphics[width=\linewidth]{images/map_sanfrancisco.png}
%\endminipage
\end{figure*}
